Displacement field paper

### 2016-09-13 ms_caf/ms_caf.tex
haoran

\begin{abstract}
$N$-body algorithms are computation-light but memory-heavy imbalanced problems in supercomputing. Traditional $N$-body simulation algorithms use at least 24-byte per particle in memory, where six 4-byte single precision real numbers keep track of each particle's phase space coordinate. This is usually the bottle neck of scaling-up of a $N$-body problem on modern supercomputers. Here we present a parallel, memory-lite Particle-Mesh based $N$-body algorithm, in which each particle can occupy as low as 9-byte memory. This is accomplished by storing only particles' relative location respect to a meshed grid, and their velocity relative to a predictable distribution function. The remaining information is given by the ordering of particles in memory space, and a complimentary density field, which is negligible in memory. Our numerical experiments show that this algorithm provides acceptable accuracy compared to traditional algorithm in cosmological simulations.

@@

\section{Introduction}
$N$-body simulation is a powerful tool to solve the highly nonlinear dynamic problems \citep{1988csup.book.....H}. It is widely used in cosmology to model the formation and evolution of the large scale structure (LSS). With the fast development of parallel supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) $N$-body particles. To date the largest $N$-body simulation in application is the ``TianNu'' simulation run on the TianHe-2 supercomputer by cosmological simulation code {\tt CUBEP3M} \citep{2013MNRAS.436..540H}. It uses nearly $3\times 10^{12}$ particles to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the cosmic age. However {\tt CUBEP3M} and TianNu did not fully use TianHe-2's computing power -- 87\% of the computing power on TianHe-2 are given by the Intel Xeon-Phi coprocessors while they occupy only 27\% of the total memory. Although we accomplished preliminary works on developing native and offload modes of {\tt CUBEP3M} on Xeon-Phi many integrated core (MIC) architecture, the memory problem is still the bottle neck of scaling up the $N$-body problem on the heterogeneous architectures. Lower the memory usage per particle will solve this problem, and will also help the $N$-body problem development on GPUs, which are even more massive parallelized and memory limited. In this paper we present a $N$-body algorithm which uses as low as 9 byte per particle (here after we use ``bpp'' referring to ``byte per particle'') and show that it can give accurate cosmological LSS simulations. The algorithm is presented in \S 2, and a comparison between this method and traditional method is shown in \S 3. Discussions and conclusions are in \S 4.

>>>>

\begin{abstract}
Cosmological large scale structure $N$-body simulations are computation-light but memory-heavy imbalanced problems in supercomputing. Traditional $N$-body simulation algorithms use at least 24-byte per particle in memory, where six 4-byte single precision real numbers keep track of each particle's phase space coordinate. This is usually the bottleneck of scaling-up of a $N$-body problem on modern supercomputers whose core/memory ratio is high. Here we present a parallel, memory-lite Particle-Mesh based $N$-body code, in which each particle can occupy as low as 6-byte memory. This is accomplished by storing only particles' relative location respect to a meshed grid, and their velocity relative to a predictable distribution function. The remaining information is given by the ordering of particles in memory space, and a complimentary density field, which is negligible in memory. Our numerical experiments show that this algorithm provides acceptable accuracy compared to traditional algorithm in cosmological simulations.

@@

\section{Introduction}
$N$-body simulation is a powerful tool to solve the highly nonlinear dynamic problems \citep{1988csup.book.....H}. It is widely used in cosmology to model the formation and evolution of the large scale structure (LSS). With the fast development of parallel supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) $N$-body particles. To date the largest $N$-body simulation in application is the ``TianNu'' simulation run on the TianHe-2 supercomputer by cosmological simulation code {\tt CUBEP3M} \citep{2013MNRAS.436..540H}. It uses nearly $3\times 10^{12}$ particles to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the cosmic age. However {\tt CUBEP3M} and TianNu did not fully use TianHe-2's computing power -- 87\% of the computing power on TianHe-2 are given by the Intel Xeon-Phi coprocessors while they occupy only 27\% of the total memory. Although we accomplished preliminary works on developing native and offload modes of {\tt CUBEP3M} on Xeon-Phi many integrated core (MIC) architecture, the memory problem is still the bottle neck of scaling up the $N$-body problem on the heterogeneous architectures. Lower the memory usage per particle will solve this problem, and will also help the $N$-body problem development on GPUs, which are even more massive parallelized and memory limited. In this paper we present a $N$-body algorithm which uses as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle'') and show that it can give accurate cosmological LSS simulations. The algorithm is presented in \S 2, and a comparison between this method and traditional method is shown in \S 3. Discussions and conclusions are in \S 4.

### 2016-09-13 ms_disp1/ms_disp1.tex
haoran

Wiener filter.

We try to decompose $\delta_L$ into two components,
\begin{equation}\label{eq.decompose}
    \delta_L=r\delta_R+\delta_N,
\end{equation}
where $r\delta_R$ is completely correlated with $\delta_L$, and the uncorrelated part $\delta_N$ is generated by the nonlinear evolution. To do this, we first correlate equation (\ref{eq.decompose}) with $\delta_R$,
\begin{equation}\label{eq.correlate}
    \left\langle\delta_L\delta_R\right\rangle
    =r\left\langle\delta_R\delta_R\right\rangle
    +\left\langle\delta_N\delta_R\right\rangle,
\end{equation}
and vanish the second term in the right-hand-side. Then the $r$ is given by $\left\langle\delta_L\delta_R\right\rangle / \left\langle\delta_R\delta_R\right\rangle$ and mode-coupling term $\delta_N=\delta_L-r\delta_R$. We next square equation (\ref{eq.decompose}) and gives
\begin{equation}\label{eq.square}
    \left\langle\delta_L\delta_L\right\rangle
    =r^2\left\langle\delta_R\delta_R\right\rangle
    +\left\langle\delta_N\delta_N\right\rangle
    +2r\left\langle\delta_R\delta_N\right\rangle,
\end{equation}
and the third term of right-hand-side also vanishes. By defining
\begin{equation}
    \left\langle\delta_L\delta_L\right\rangle
    =b^2\left\langle\delta_R\delta_R\right\rangle,
\end{equation}
we write
\begin{equation}
    \left\langle\delta_N\delta_N\right\rangle=(b^2-r^2)\left\langle\delta_R\delta_R\right\rangle
\end{equation}
and a Wiener filter can be constructed as
\begin{equation}
    W=\frac{r^2\left\langle\delta_R\delta_R\right\rangle}
    {r^2\left\langle\delta_R\delta_R\right\rangle+\left\langle\delta_N\delta_N\right\rangle}
    =\frac{r^2}{b^2}.
\end{equation}

From equation (\ref{eq.decompose}), multiplying $W$, we obtain the optimal reconstructed linear density field by
\begin{equation}
    \tilde\delta_L=rW\delta_R=\frac{r^3}{b^2}\delta_R.
\end{equation}
From equation (\ref{eq.square}), multiplying $W^2$, the reconstructed linear power spectrum is giving by
\begin{equation}
\begin{aligned}
    \tilde{P}&=W^2\left\langle\delta_L\delta_L\right\rangle  \\
    &=r^2W^2\left\langle\delta_R\delta_R\right\rangle +W^2\left\langle\delta_N\delta_N\right\rangle  \\
    &=b^2W^2\left\langle\delta_R\delta_R\right\rangle  \\
    &=r^4b^{-2}\left\langle\delta_R\delta_R\right\rangle.
\end{aligned}
\end{equation}

>>>>

Wiener filter.
Define
\begin{equation}
    r\equiv \frac{\left\langle\delta_L\delta_R\right\rangle}
    {\sqrt{\left\langle\delta_L\delta_L\right\rangle
    \left\langle\delta_R\delta_R\right\rangle}}
\end{equation}
and
\begin{equation}
    b^2\equiv \frac{\left\langle\delta_R\delta_R\right\rangle}
    {\left\langle\delta_L\delta_L\right\rangle}.
\end{equation}

We try to decompose $\delta_L$ into two components,
\begin{equation}\label{eq.decompose}
    \delta_L=r'\delta_R+\delta_N,
\end{equation}
where $r'\delta_R$ is completely correlated with $\delta_L$, and the uncorrelated part $\delta_N$ is generated by the nonlinear evolution. To do this, we first correlate equation (\ref{eq.decompose}) with $\delta_R$,
\begin{equation}\label{eq.correlate}
    \left\langle\delta_L\delta_R\right\rangle
    =r'\left\langle\delta_R\delta_R\right\rangle
    +\left\langle\delta_N\delta_R\right\rangle,
\end{equation}
and vanish the second term in the right-hand-side. Then
\begin{equation}
    r'=\frac{\left\langle\delta_L\delta_R\right\rangle}
    {\left\langle\delta_R\delta_R\right\rangle}=\frac{r}{b}.
\end{equation}
Then the mode-coupling term 
\begin{equation}
    \delta_N=\delta_L-\frac{r}{b}\delta_R.
\end{equation}
We next square equation (\ref{eq.decompose}) and gives
\begin{equation}\label{eq.square}
    \left\langle\delta_L\delta_L\right\rangle
    =\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle
    +\left\langle\delta_N\delta_N\right\rangle
    +2\frac{r}{b}\left\langle\delta_R\delta_N\right\rangle,
\end{equation}
and the third term of right-hand-side also vanishes. As a result,
\begin{equation}
    \left\langle\delta_N\delta_N\right\rangle
    =\frac{1-r^2}{b^2}
    \left\langle\delta_R\delta_R\right\rangle
\end{equation}
and a Wiener filter can be constructed as
\begin{equation}
    W=\frac{r^2}{r^2+1-r^2}=r^2.
\end{equation}

From equation (\ref{eq.decompose}), multiplying $W$, we obtain the optimal reconstructed linear density field by
\begin{equation}
    \tilde\delta_L=W\frac{r}{b}\delta_R=\frac{r^3}{b}\delta_R.
\end{equation}
From equation (\ref{eq.square}), multiplying $W^2$, the reconstructed linear power spectrum is giving by
\begin{equation}
\begin{aligned}
    \tilde{P}&=W^2\left\langle\delta_L\delta_L\right\rangle  \\
    &=r^2\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle +r^2\left\langle\delta_N\delta_N\right\rangle  \\
    &=\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle.
\end{aligned}
\end{equation}
